Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-10 paper/paper.tex
major edits to text in paper Tested everything on quipu.

Studies have shown that the cell division process is capable of
preventing minicelling for fairly perturbed cell
shapes\cite{touhami2006temperature}. Varma et al. studied one
particular example of a perturbed shape, a three-pronged tube, and
found both experimental and simulation results\cite{varma2008min} that
show oscillation. These oscillations seem to seek out the extreme
poles in the cell shape\cite{corbin2002exploring}
\cite{juarez2010changes}, and effectively prevent minicelling in most
cases. However, Mannik et al. have also shown that there are
limitations to this mechanism\cite{mannik2010bacteria}
\cite{mannik2009bacterial}. One instance of this behavior is
observable when the cell has been forced into a flattened, irregular
shape. Oscillations in this sort of cell shape are far from the
regular sorts of oscillations seen in the computational work done by
Huang et al\cite{huang2003dynamic}. We seek to apply a computational
model to explore the more extreme or irregular cell shapes, in order
to determine the limits of the steady cell oscillations seen by both
theory and experiment.

A significant amount of work has been done to develop protein reaction
and diffusion models that exhibit accurate macroscopic dynamics of the
MinD protein system. Early models involve free proteins that affect
each others' rates of diffusion and membrane attachment, but do not
combine into compound states~\cite{meinhardt2001pattern}.  In 2003
Huang improved upon this work with a simple simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in
cylindrical cells~\cite{huang2003dynamic}. In this model cytoplasmic
MinD is more likely to attach to the membrane when MinD is already
clustered there (following observed non-linear attachement of minD on
the cell membrane), and is stationary once attached.  A number of
studies have used an approach similar in that they do not rely on the
ability of MinD to move along the walls and
cluster~\cite{kruse2007experimentalist, meinhardt2001pattern,
  drew2005polymerization, fange2006noise, kerr2006division}, while
studies have been made as well of models which rely on MinD mobility
and attraction on the cell membrane~\cite{kruse2002dynamic,
  howard2005cellular}.  Variations of the Huang 2003 model that
stochasticly account for variations of molecular interaction
\cite{fange2006noise} and monte-carlo simulations that implement this
stochastic version of Huang's mean field reaction rates confirm the
major results obtained by Huang's model, and more successfully predict
experimentally observed oscillations in round cell
phenotypes~\cite{drew2005polymerization, fange2006noise,
  huang2004min}. Biochemical models of broader scope have also been
used to study the MinD system and show consistent
results.\cite{arjunan2010new}.  However, in general, the results of
the stochastic and monte-carlo simulations are similar to those given
by Huang's mean field results.

Previous studies have been made of the Min system's associtation with
the cell membrane
\cite{hsieh2010direct}\cite{mileykovskaya2003effects}.  Studies have
shown as well that MinD binds preferentially to regions enriched with
cadiolipin, an anionic phospholipid that collects on regions of high
negative curvature. This mechanism has been incorporated into other
models.\cite{drew2005polymerization,cytrynbaum2007multistranded,renner2012mind,renner2012mind}
However, this mechanism of combined clustering, phospholipids and MinD
has not been observed in real cells. \cite{halatek2012highly}

@@

\section{Model, Methods, and Pill Shaped Cell}
We implement the model of Huang \emph{et al.}  for the behavior of the
MinD and MinE proteins inside the cell~\cite{huang2003dynamic}.
Figure~\ref{fig:reactions} shows the reaction process.  The
cytoplasmic MinD:ADP complex undergoes nucleotide exchange and is
changed into the MinD:ATP complex.  This will naturally diffuse
towards and attach to the cell membrane.  A cytoplasmic MinE will
attach to the wall bound MinD:ATP complex and after a time will
activate ATP hydrolosis.  This breaks up the complex, releasing MinE,
phosphate, and MinD:ADP back into the cytoplasm.  The MinD:ADP will undergo
nucleotide exchange and begin again the cyclic process.  This model is
defined by a set of five reaction-diffusion equations:

@@

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular size and shape on the protein
oscillations, so we follow Huang\cite{huang2003dynamic} and do not
deviate from the wild type values used in the cited work (see values
below).  Huang's simulations use total MinD and MinE concentrations of
$1,000/\mu m$ and $350/\mu m$, respectively, in a cylindrical cell of
radius $0.5\mu m$, and in our (non-cylindrical) cells we use the same
number of proteins per unit volume.  These concentration values are
1273 $\mu m^{-3}$ and 446 $\mu m^{-3}$, respectively.
\begin{gather*}
  \mathcal{D}_D = \mathcal{D}_{E} = 2.5\micron^2/\text{sec}\\
  k_D^{\textrm{ADP $\rightarrow$ ATP}} = 1/\textrm{sec,  }
  k_D = 0.025 \micron /\textrm{sec}\\
  k_{dD} = 0.0015 \micron^3/ \textrm{sec,  }
  k_{de} = 0.7/\textrm{sec}\\
  k_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}
A 3D grid is then constructed in cartesian coordinates, with a grid
spacing of .05\micron.

In order to further our understanding of the results of various
simulation approaches to this system, we have performed both a basic
finite analysis, deterministic simulation that is spatially and
temporally discrete, and a stochastic analysis that is spatially
discrete but continuous in time.  Our stochastic simulation method
follows the work of Kraus~\cite{kraus1996crosstalk} which in turn
follows a method introduced by Gillespie~\cite{gillespie1977exact}.

We mean to investigate the geometric limits of the Min system
oscillations as observed by Mannik \emph{et
  al.}~\cite{mannik2012robustness}, so we model the Min system in a
variety of cell shapes and sizes.  Here we present a selection of
those, first naturally occuring pill-shaped cells and then a
number of flattened out shapes which reflect the experiments of Mannik
\emph{et al.}, in which bacteria are confined within a thin slit of
height $0.25\micron$ ~\cite{mannik2012robustness}. Viewed from the top
down the cells will have the shapes described below and viewed from
the side they have at their edges a semicircular protrusion (one may
imagine the edges of a pancake). Central to our observations are the
effects of cell size and so we show expanded cells that have the same
thickness but increased dimensions in the two planar directions.  We
present two shapes that are designed to replicate those published by
Mannik, equilateral and iscosolese triangle shapes, and a 'V' shape of
our own creation in order to best illustrate the effect of cell size
on oscillation stability.

@@

Kubitschek has shown in multiple experiments that at the time of cell
division cells have a volume that is within a range of roughly
$1\micron^3$ to $2\micron^3$~\cite{kubitschek1990cell,
  kubitschek1968linear}.  We follow Huang's
simulations\cite{huang2003dynamic} and Mannik's experiments and model
cells that are slightly larger than this range.

\begin{figure*}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/image-plot}
  %\includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/single-image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/single-image-plot}
  \caption{\fixme{We should figure out units and use the same levels
      as in the main plot.  I also want to figure out why the two
      plots look different, and make them the same size.  Can we just
      do the gaussian} Contour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic simulation and the lower shows results from the
    stochastic simulation.  The cells pictured are $4\micron$ in
    length, measured from end to end.  The order of frames is such
    that individual MinD proteins begin at the bottom of the plot (in
    the MinD:ATP state in the cytoplasm), and progress upward until
    they reach the MinE:MinD:ATP membrane-bound complex.  At that
    point, they will spontaneously dissociate into cytoplasmic MinE
    (the top row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.  The
    stochastic densities are smeared out over space using a guassians
    approximation developed by Zhang et all~\cite{zhang2007gaussian}}.
  \label{image-p}
\end{figure*}

We begin with the naturally occuring pill cell shape.  We peice this
shape together as two hemispherical endcaps attached on either end of
a cylinder.  This shape follows the early simulations of Huang
\emph{et al.} but differs in that we have added the end caps for a
more natural shape, expecting similar results.  We test cells of
radius $.5\micron$ and of lengths $2\micron$, $3\micron$, and
$5\micron$ measured from the tip of one end cap to the tip of the
other. In each case, we observe in the deterministic simulations a
quick establishement (within a few periods) of extremely regular
oscillatory behavoir that lasts indefinitely.  The stochastic results
show oscillitions that are similarly regular in time, with periods
that nearly match those of the deterministic (they differ roughly by a
factor of .05). The stochastic results also show a slight variation in
the height of the density maxima, but the peaks are still very well
defined. The period increases with cell length, as shown in
Table~\ref{tab:pill-periods}, in agreement with the results of Huang
\emph{et al.}.  The stochastic simulations show a similar pattern.
The MinD maxima occur at the ends of the cell in a back and forth
manner, and with the same approximate period.

@@

This simple cell shape is a good starting point for observing in
detail the dynamic interaction between the different stages of protein
that lead to their qualitatively distinct behavoir. Figure
\ref{image-p} shows the process in a series of frames taken from
simulation animations for both the deterministic and stochastic
simulations.  We have 'smeared out' the stochastic concentrations in a
manner meant to reproduce the images shown by diffraction limited
flourescence microscopy.  We do so using the two dimensional gaussian
approximation developed by Zhang et all~\cite{zhang2007gaussian}.  In
this approximation we use a numerical aperture value of 1.3, which is
the same as used by mannik experimentally, and a wavelength of
$509nm$, typical of green flourescent microscopy.  Each frame is 2.5
seconds ahead of the last, and each image shows a the concentration of
a given state of protein summed over the coordinate orthogonal to the
page.  The color scale for each protein state is set according to the
maximum values observed.

At $t=0$ in Figs.~\ref{image-p}, the cytoplasmic MinE is concentrated
in the lower portion of the cell and is diffusing upward, since it
will react with and stick to wall-bound MinD that is concentrated on
the membrane in the upper portion of the cell.  At the boundary
between the cytoplasmic MinE and the membrane-bound MinD, there is a
ring of the membrane-bound MinE:MinD complex, which is moving towards
the upper end of the cell, removing the MinD from the membrane as it
goes.  The MinD is removed from the membrane in its MinD:ADP form and
diffuses freely for a time before spontenously undergowing nucleotide
exchange and shifting back into the MinD:ATP that is able to attach to
the membrane.  Diffusion away from the end of the cell will bring the
protein into the MinE ring, where reactions will prevent it from
collecting for any significant time on the membrane.  Diffusion
further towards the upper end, however, while the MinE has not yet
collected there, will allow MinD:ATP to accumulate with the other
MinD:ATP, peaking at around 10 seconds in Figure~\ref{image-p}.  As
the MinE ring approaches the end of the cell, seen in the seconds
following 10, the only freely diffusing MinD:ADP that 'escapes' the
MinE will be that which has diffused away from the the end of the cell
and past the MinE ring towards the center.

At 15 seconds, the membrane-bound MinD:ATP has been essentially
removed from the top end of the cell, and MinD:ATP has begun to bind
to the membrane in the the lower half of the cell.  At this stage,
there is a high concentration of cytoplasmic MinE at the top of the
cell, and by 16 seconds we begin to see the formation of a MinE ring
just below the center of the cell.  At 18 seconds, the cell has
reached its original state (reversed directionally), with MinD:ATP
bound to the membrane on the lower third of the cell, a high
concentration of cytoplasmic MinE in the upper half of the cell, and a
MinE ring pushing downward on the membrane-bound MinD:ATP.

>>>>

Studies have shown that the cell division process is capable of
preventing minicelling for fairly perturbed cell
shapes\cite{touhami2006temperature}. Varma et al. studied one
particular example of a perturbed shape, a three-pronged tube, and
found both experimental and simulation results\cite{varma2008min} that
show oscillation. These oscillations seem to seek out the extreme
poles in the cell shape\cite{corbin2002exploring}
\cite{juarez2010changes}, and effectively prevent minicelling in most
cases. However, Mannik et al. have also shown that there are
limitations to this mechanism\cite{mannik2010bacteria}
\cite{mannik2009bacterial}. One instance of this behavior is
observable when the cell has been forced into a flattened, irregular
shape. Oscillations in this sort of cell shape are far from the
regular sorts of oscillations seen in the computational work done by
Huang et al\cite{huang2003dynamic}, who simulate more regular
shapes. Thus there is an opportunity here to test previously used
simulation models against more extreme experimental cases than have
been seen thus far.  In this paper we apply a deterministic method and
a stochastic method of simulation and compare them against Mannik's
experimental findings.

A significant amount of work has been done to develop protein reaction
and diffusion models that exhibit accurate macroscopic dynamics of the
MinD protein system. Early models involve free proteins that affect
each others' rates of diffusion and membrane attachment, but do not
combine into compound states~\cite{meinhardt2001pattern}.  In 2003
Huang improved upon this work with a simple and very successful
simulation model based on MinD-MinE combination, ATPase hydrolysis,
and MinD membrane attachment that exhibits accurate MinD oscillations
in cylindrical cells~\cite{huang2003dynamic}. In this model
cytoplasmic MinD is more likely to attach to the membrane when MinD is
already clustered there (following observed non-linear attachement of
minD on the cell membrane), and is stationary once attached.  A number
of studies have used an approach similar in that they do not rely on
the ability of MinD to move along the walls and
cluster~\cite{kruse2007experimentalist, meinhardt2001pattern,
  drew2005polymerization, fange2006noise, kerr2006division}, while
studies have been made as well of models which rely on MinD mobility
and attraction on the cell membrane~\cite{kruse2002dynamic,
  howard2005cellular}.  Experimental studies of the Min system's
association with the cell membrane have also been
made~\cite{hsieh2010direct}\cite{mileykovskaya2003effects}.

Variations of the Huang 2003 model that stochasticly account for
variations of molecular interaction \cite{fange2006noise} and
monte-carlo simulations that implement this stochastic version of
Huang's mean field reaction rates confirm the major results obtained
by Huang's model, and more successfully predict experimentally
observed oscillations in round cell
phenotypes~\cite{drew2005polymerization, fange2006noise,
  huang2004min}. Biochemical models of broader scope have also been
used to study the MinD system and show consistent
results.\cite{arjunan2010new}.  However, in general, the results of
the stochastic and monte-carlo simulations have so far yeilded similar
results to those given by Huang's original, deterministic method of
simulation.\fixme{Look at the papers again to verify and this and also
  figure out and cite who says that the stochastic and the
  deterministic methods of simulation have given similar results.
  Saying more about these discussions is needed here, I think.
  Something to the effect that until now there's no compelling reason
  to avoid the deterministic simulations.}

\fixme{Following paragraph is awkward, and I don't think it's needed.}
Studies have shown as well that MinD binds preferentially to regions
enriched with cadiolipin, an anionic phospholipid that collects on
regions of high negative curvature. This mechanism has been
incorporated into other
models.\cite{drew2005polymerization,cytrynbaum2007multistranded,renner2012mind,renner2012mind}
However, this mechanism of combined clustering, phospholipids and MinD
has not been observed in real cells. \cite{halatek2012highly}

@@

\section{Model, Methods, and Cell Shapes}
We implement the reaction-diffusion model of Huang \emph{et
  al.}~\cite{huang2003dynamic}.  Figure~\ref{fig:reactions} shows the
reaction process.  The cytoplasmic MinD:ADP complex undergoes
nucleotide exchange and is changed into the MinD:ATP complex.  This
will naturally diffuse and attach to the cell membrane.  A cytoplasmic
MinE will attach to the wall bound MinD:ATP complex and after a time
will activate ATP hydrolosis.  This breaks up the complex, releasing
MinE, phosphate, and MinD:ADP back into the cytoplasm.  The MinD:ADP
will undergo nucleotide exchange and begin again the cyclic process.
The model is defined by a set of five reaction-diffusion equations:

@@

Our diffusion and reaction rates are shown below.  We are interested
primarily in the effect of cellular size and shape on the protein
oscillations, so we follow Huang\cite{huang2003dynamic} and do not
deviate from the wild type values used in the cited work (see values
below).  Huang's simulations use total MinD and MinE concentrations of
$1,000/\mu m$ and $350/\mu m$, respectively, in a cylindrical cell of
radius $0.5\mu m$, and in our (non-cylindrical) cells we use the same
number of proteins per unit volume.  These concentration values are
1273 $\mu m^{-3}$ and 446 $\mu m^{-3}$, respectively. \fixme{It's been
  a while so might want to check this to make sure it's true.}
\begin{gather*}
  \mathcal{D}_D = \mathcal{D}_{E} = 2.5\micron^2/\text{sec}\\
  k_D^{\textrm{ADP $\rightarrow$ ATP}} = 1/\textrm{sec,  }
  k_D = 0.025 \micron /\textrm{sec}\\
  k_{dD} = 0.0015 \micron^3/ \textrm{sec,  }
  k_{de} = 0.7/\textrm{sec}\\
  k_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}
A 3D grid is then constructed in cartesian coordinates, with a grid
spacing of .05\micron.

We have performed both a basic finite analysis, deterministic
simulation that is spatially and temporally discrete, and a stochastic
analysis that is spatially discrete but continuous in time.  Our
stochastic simulation method follows the work of
Kraus~\cite{kraus1996crosstalk} which in turn follows a method
introduced by Gillespie~\cite{gillespie1977exact}.

We mean to investigate the geometric limits of the Min system
oscillations as observed by Mannik \emph{et
  al.}~\cite{mannik2012robustness}, so we model the Min system in a
variety of cell shapes and sizes.  Here we present a selection of
these, first naturally occuring pill-shaped cells and then a number of
flattened out shapes which reflect the experiments of Mannik \emph{et
  al.}, in which bacteria are confined within a thin slit of height
$0.25\micron$ ~\cite{mannik2012robustness}. Viewed from the top down
the cells will show the shapes described below and viewed from the
side they show at their edges a semicircular protrusion (one may
imagine the shape of a pancake). We have simulated and analyzed a
number of these flattened shapes in addition to those shown below,
including equilateral and iscosolese triangle shapes, shapes shaped
like greater than signs, '$>$', and four-pronged 'starfish' shapes.
For all of these we have varied the cell size, expanding the two
dimensions that are tangent to the flattened surface from very small,
'steady state' scales (which we discuss below) to scales twice that of
the published Mannik shapes.  Our conclusions are based on all of this
data, but we present here specific, illustrative examples: two
flattened shapes that are designed to replicate those published by
Mannik and two 'stadium' shapes that respectively have the same
quadrupole moments of the two Mannik shapes.

@@

\fixme{Is the following Kubitscheck paragraph needed?}
Kubitschek has shown in multiple experiments that at the time of cell
division cells have a volume that is within a range of roughly
$1\micron^3$ to $2\micron^3$~\cite{kubitschek1990cell,
  kubitschek1968linear}.  We follow Huang's
simulations\cite{huang2003dynamic} and Mannik's experiments and model
cells that are slightly larger than this range.

\section{Naturally Occuring Pill Shaped Cells}
\begin{figure*}
  %\includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-exact/plots/single-image-plot}
  \includegraphics[width=\textwidth]{../data/shape-p/3_00-0_50-0_00-0_00-15_00-full_array/plots/single-image-plot}
  \caption{\fixme{We should figure out units and use the same levels
      as in the main plot.  I also want to figure out why the two
      plots look different, and make them the same size.  Can we just
      do the gaussian} Contour plot images of the concentration of each protein
    species in a natural pill-shaped bacterium at regular intervals in
    time (one second intervals), with darker regions indicating higher
    concentration. The upper plots shows results from the
    deterministic simulation and the lower shows results from the
    stochastic simulation.  The cells pictured are $4\micron$ in
    length, measured from end to end.  The order of frames is such
    that individual MinD proteins begin at the bottom of the plot (in
    the MinD:ATP state in the cytoplasm), and progress upward until
    they reach the MinE:MinD:ATP membrane-bound complex.  At that
    point, they will spontaneously dissociate into cytoplasmic MinE
    (the top row) and the starting state of cytoplasmic MinD:ADP.  The
    densities plotted are integrated along the axis orthogonal to the
    page, and the color scale is chosen separately for each species
    with black as the maximum value over space and time.  The
    stochastic densities are smeared out over space using a guassians
    approximation developed by Zhang et all~\cite{zhang2007gaussian}}.
  \label{image-p}
\end{figure*}

We begin with the naturally occuring pill cell shape.  We peice this
shape together as two hemispherical endcaps attached on either end of
a cylinder.  This shape follows the early simulations of Huang
\emph{et al.} but differs in that we have added the end caps for a
more natural shape, expecting similar results.

This simple cell shape is a good starting point for observing in
detail the dynamic interaction between the different stages of protein
that lead to their qualitatively distinct behavoir. Figure
\ref{image-p} shows the process in a series of frames taken from
simulation animations for both the deterministic and stochastic
simulations.  We have 'smeared out' the stochastic concentrations in a
manner meant to reproduce the images shown by diffraction limited
flourescence microscopy.  We do so using the two dimensional gaussian
approximation developed by Zhang et all~\cite{zhang2007gaussian}.  In
this approximation we use a numerical aperture value of 1.3, which is
the same as used by mannik experimentally, and a wavelength of
$650nm$.  Each frame is 2.5 seconds ahead of the last, and each image
shows the concentration of a given state of protein (of the five
described in the reaction model) summed over the coordinate orthogonal
to the page.  The color scale for each protein state is set according
to the maximum values observed.

Figure~\ref{image-p} begins about 300 seconds into the simulation and
shows a typical period of oscillation that in the pill shape repeates
itself with a high degree of regularity, for both the deterministic
and stochastic simulations.  At $t=0$ the cytoplasmic MinE is
concentrated in the upper portion of the cell and is diffusing
downward, where it will react with and stick to wall-bound MinD that
is concentrated on the membrane in the lower portion of the cell.
This process is in fact already well on its way, as seen by the
significant concentration of wall bound MinD:MinE which is creeping
down to the lower corner, removing the MinD from the membrane as it
goes.  The removed MinD (cytoplasmic MinD:ADP) diffuses freely for a
time before spontenously undergowing nucleotide exchange and shifting
back into the MinD:ATP that is able to attach to the membrane.
Diffusing a small distance away from the bottom of the cell will bring
the protein into the MinE collection, where reactions will prevent it
from collecting for any significant time on the membrane.  Diffusing
further towards the upper end, however, while the MinE has not yet
collected there, will allow MinD:ATP to accumulate with the other
MinD:ATP, which is seen at seconds 10 through 15 in
Figure~\ref{image-p}.  Eventually the MinE creeps all the way down to
the bottom of the cell where it slowly runs out of wall bound MinD to
react with, before diffusing back up and starting again on the upper
portion of the cell. This is seen in seconds 15 through 20 in the
figure.  \fixme{There is another paragraph after this in the latex
  that I commented out because I thought we might be spending too much
  time on the description}

%%  At 15 seconds, the membrane-bound MinD:ATP has
%% been essentially removed from the top end of the cell, and MinD:ATP
%% has begun to bind to the membrane in the the lower half of the cell.
%% At this stage, there is a high concentration of cytoplasmic MinE at
%% the top of the cell, and by 16 seconds we begin to see the formation
%% of a MinE ring just below the center of the cell.  At 18 seconds, the
%% cell has reached its original state (reversed directionally), with
%% MinD:ATP bound to the membrane on the lower third of the cell, a high
%% concentration of cytoplasmic MinE in the upper half of the cell, and a
%% MinE ring pushing downward on the membrane-bound MinD:ATP.

We test cells of radius $.5\micron$ and of lengths $2\micron$,
$3\micron$, and $5\micron$ measured from the tip of one end cap to the
tip of the other. In each case, we observe in the deterministic
simulations a quick establishement (within a few periods) of extremely
regular oscillatory behavoir that lasts indefinitely.  The stochastic
results show oscillitions that are similarly regular in time, with
periods that nearly match those of the deterministic (they differ
roughly by a factor of .05)\fixme{does this square with the
  correlation data periods?}.

While the stochastic results show periods that seem to have a high
degree of regularity as well as similarity to the deterministic
results, they do differ in the precise locations of the maxima.  While
the deterministic simulations show maxima that are every time
concentrated in the center of one of the cell ends, with a symmetric
smear of protein going out laterally to either side, the stochastic
simulations show maxima that, while always located somehwere in the
end region of the cell, occur at different positions within this
region.  This is reflected in Figure~\ref{image-p}, which shows
periods that are the same and qualitative behavoir that is
approximately the same, but also that there is variation in the
details of the stochastic simulation maxima location.

The period increases with cell length, as shown in
Table~\ref{tab:pill-periods}, in agreement with the results of Huang
\emph{et al.}.  The stochastic simulations show a similar pattern.
The MinD maxima occur at the ends of the cell in a back and forth
manner, and with the same approximate period.

@@
